### 2024-02-27 presentation/parts/introduction.tex
WIP

\begin{frame}
    \frametitle{Medical dialogue}
    \begin{figure}
        \centering
        \includegraphics[width=0.7\paperwidth]{figures/speech_bubble.pdf}
    \end{figure}

    \note[item]{Healthcare sector is a complex system}
\end{frame}

\begin{frame}
    \frametitle{Errors in medical dialogue}
    \begin{itemize}
        \item<1,2>Communication is everywhere in healthcare. 
        \item <1,2>It is complex, involving multiple participants, different contexts, and different purposes.
        \vspace{1em}
        \item <2> \highlight{Adverse events}: Failure of communication is a leading cause of medical error contributing to two out of three adverse events \cite{starmer_changes_2014}.
        \item <2> \highlight{Preventability}: A considerable fraction of all hospital admissions had preventable adverse outcomes (9\% to 16.6\% in AU, NZ, UK, DK) \cite{carver_medical_2024}.
    \end{itemize}
\end{frame}

% \begin{frame}
%     \frametitle{Types of dialogue}
%     \begin{itemize}
%         \item Domain: Specialized / General
%         \item Context: Controlled / Chaotic
%         \item Person: Nurse, doctor, midwife, caregiver, psychiatrist, insurance professional
%         \item Purpose: Triaging, diagnosis, treatment, follow-up, documentation, coding, billing
%     \end{itemize}
% \end{frame}

\begin{frame}
    \frametitle{Documenting medical encounters}
    \begin{itemize}
        \item <1,2> Documentation is a central part of healthcare.
        \item <1,2> E.g. patient records, insurance claims, billing, research, training, legal purposes.
        \vspace{1em}
        \item <2> \highlight{Time-consuming}: Physicians spend 34-37\% of their time on documentation \cite{joukes_time_2018, tipping_where_2010, sinsky_allocation_2016}\footnote{Ambulatory care across four specialties in four states and tertiary care at an academic medical center.}.
        \item <2> \highlight{Varying quality}: Discharge summaries rarely meet all timeline, transmission, and content criteria. \cite{horwitz_comprehensive_2013}\footnote{Outpatient visits, Yale-New Haven Hospital.}
    \end{itemize}

    \note[item]{Documentation is essential for a number of purposes but is very time-consuming and imperfect}
\end{frame}

\begin{frame}
    \frametitle{How might machine learning help?}
    \begin{itemize}
        \item <1-> \highlight{Assist} with documentation.
        \item <1-> \highlight{Augment} communication.
        \item <1-> \highlight{Improve} decision-making.
        \vspace{1em}
        \item <2-> \highlight{Reduce} number of impact of medical errors and adverse events.
        \item <2-> \highlight{Free up} time spent on documentation for patient care.
    \end{itemize}
\end{frame}

>>>>

% \begin{frame}
%     \frametitle{Medical dialogue}
%     \begin{figure}
%         \centering
%         \includegraphics[width=0.7\paperwidth]{figures/speech_bubble.pdf}
%     \end{figure}

%     \note[item]{Healthcare sector is a complex system}
% \end{frame}
% \begin{frame}
%     \frametitle{Medical dialogue}
%     \begin{figure}
%         \centering
%         \includegraphics[width=0.95\textheight]{figures/corti_sketch_conversations.png}
%     \end{figure}
% \end{frame}

\begin{frame}[t]
    \frametitle{Medical dialogue}
    \begin{columns}[t]
        \begin{column}{0.4\textwidth}
            \begin{itemize}
                \item <2> General practitioner
                \item <2> Nurse
                \item <2> Midwife
                \item <2> Emergency medical dispatcher
                \item <2> Paramedic
                \item <2> Emergency room
                \item <2> Health insurance
            \end{itemize}
        \end{column}
        \begin{column}{0.6\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.9\textwidth]{figures/corti_sketch_conversations.png}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Errors in medical dialogue}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item <1-> Communication is everywhere in healthcare. 
                \item <1-> It is complex, involving multiple participants, different contexts, and different purposes.
                \vspace{1em}
                \item <2-> \highlight{Adverse events}: Failure of communication is a leading cause of medical error contributing to two out of three adverse events \cite{starmer_changes_2014}.
                \item <2-> \highlight{Preventability}: A considerable fraction of all hospital admissions have preventable adverse outcomes\footnote<2>{9\% to 16.6\% in AU, NZ, UK, DK.} \cite{carver_medical_2024}.
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.95\textwidth]{figures/corti_sketch_emergency.png}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Documenting medical encounters}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item <1,2> Documentation is a central part of healthcare.
                \item <1,2> E.g. patient records, insurance claims, billing, research, training, legal purposes.
                \vspace{1em}
                \item <2> \highlight{Time-consuming}: Physicians spend 34-37\% of their time on documentation \cite{joukes_time_2018, tipping_where_2010, sinsky_allocation_2016}\footnote<2>{Ambulatory care across four specialties in four states and tertiary care at an academic medical center.}.
                \item <2> \highlight{Varying quality}: Discharge summaries almost never meet \emph{all} timeline, transmission, and content criteria. \cite{horwitz_comprehensive_2013}\footnote<2>{Outpatient visits, Yale-New Haven Hospital.}
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=0.95\textwidth]{figures/corti_sketch_responders.png}
            \end{figure}
        \end{column}
    \end{columns}
    \note[item]{Documentation is essential for a number of purposes but is very time-consuming and imperfect}
\end{frame}

\begin{frame}
    \frametitle{How might machine learning help?}
    \begin{columns}
        \begin{column}{0.35\textwidth}
            \begin{itemize}
                \item <1-> \highlight{Assist} with documentation.
                \item <1-> \highlight{Augment} communication.
                \item <1-> \highlight{Improve} decision-making.
                \vspace{1em}
                \item <2-> \highlight{Reduce} number of impact of medical errors and adverse events.
                \item <2-> \highlight{Free up} time spent on documentation for patient care.
            \end{itemize}
        \end{column}
        \begin{column}{0.65\textwidth}
            \centering
            \includegraphics[width=0.95\textwidth]{figures/corti_sketch_office.png}
        \end{column}
    \end{columns}
\end{frame}

### 2024-02-27 presentation/parts/presentation_hierarchical.tex
WIP

\frame{
    \frametitle{Problem and Contributions}
    \begin{itemize}
        \item Deep generative models often fail at OOD detection task when using their likelihood estimate as the score function \cite{nalisnick_deep_2019} by, perhaps surprisingly, assigning \textbf{higher likelihoods} to the OOD data.
        \item Contributions:
        \begin{itemize}
            \item We provide evidence that out-of-distribution detection fails due to learned low-level features that generalize across datasets.
            \item We present a fast and fully unsupervised method for OOD detection competitive with the state-of-the-art
        \end{itemize}  
    \end{itemize}
}

@@

@@

\begin{frame}
    \frametitle{Out-of-distribution detection with hierarchical VAEs}
    \begin{columns}
        \begin{column}{0.4\textwidth}
            \begin{itemize}
                \item Generative models learn to approximate the \textbf{data distribution} $p(\xb)$.
                \item The likelihood of the model given a sample $\xb$ is a measure of how well the model \textbf{explains the data}.
                \item \textbf{Model likelihood} have long been thought of as useful for OOD detection \cite{bishop_novelty_1994}.
            \end{itemize}
        \end{column}
        \begin{column}{0.6\textwidth}
            \begin{overprint}
                \onslide<2>
                \begin{figure}
                    \centering
                    \includegraphics[width=\textwidth]{../graphics/paper_hierarchical/densities-FashionMNIST-test-MNIST-test-bpp-k-0_sohau.pdf}
                \end{figure}
            \end{overprint}
        \end{column}
    \end{columns}
\end{frame}

@@

@@

\frame{
    \frametitle{Results on diverse datasets}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{table}[t]
                \centering
                \resizebox{0.9\textwidth}{!}{%
                \begin{tabular}{llrrr}
                    \toprule
                    OOD dataset & Metric & AUROC$\uparrow$ & AUPRC$\uparrow$ & FPR80$\downarrow$ \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on CIFAR10}} \\
                    \midrule
                    SVHN          &  $LLR^{>2}$  &  0.811  &  0.837  &  0.394  \\
                    CIFAR10       &  $LLR^{>1}$  &  0.469  &  0.479  &  0.835  \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on SVHN}} \\
                    \midrule
                    CIFAR10            &  $LLR^{>1}$       &  $0.939$  &  $0.950$  &  $0.052$  \\
                    SVHN               &  $LLR^{>1}$       &  $0.489$  &  $0.484$  &  $0.799$  \\
                    \bottomrule
                \end{tabular}
                }
            \end{table}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{table}[t]
                \centering
                \resizebox{0.9\textwidth}{!}{%
                \begin{tabular}{llrrr}
                    \toprule
                    OOD dataset & Metric & AUROC$\uparrow$ & AUPRC$\uparrow$ & FPR80$\downarrow$ \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on FashionMNIST}} \\
                    \midrule
                    MNIST                    & $LLR^{>1}$             &  0.986  &  0.987  &  0.011 \\
                    notMNIST                 &  $LLR^{>1}$            &  0.998  &  0.998  &  0.000 \\
                    KMNIST                   &  $LLR^{>1}$            &  0.974  &  0.977  &  0.017 \\
                    Omniglot28x28            &  $LLR^{>2}$            &  1.000  &  1.000  &  0.000 \\
                    Omniglot28x28Inverted    &  $LLR^{>1}$            &  0.954  &  0.954  &  0.050 \\
                    SmallNORB28x28           &  $LLR^{>2}$            &  0.999  &  0.999  &  0.002 \\
                    SmallNORB28x28Inverted   &  $LLR^{>2}$            &  0.941  &  0.946  &  0.069 \\
                    FashionMNIST             &  $LLR^{>1}$            &  0.488  &  0.496  &  0.811 \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on MNIST}} \\
                    \midrule
                    FashionMNIST                   &  $LLR^{>1}$  &  $0.999$  &  $0.999$  &  $0.000$ \\
                    notMNIST                       &  $LLR^{>1}$  &  $1.000$  &  $0.999$  &  $0.000$ \\
                    KMNIST                         &  $LLR^{>1}$  &  $0.999$  &  $0.999$  &  $0.000$ \\
                    Omniglot28x28                  &  $LLR^{>1}$  &  $1.000$  &  $1.000$  &  $0.000$ \\
                    Omniglot28x28Inverted          &  $LLR^{>1}$  &  $0.944$  &  $0.953$  &  $0.057$ \\
                    SmallNORB28x28                 &  $LLR^{>1}$  &  $1.000$  &  $1.000$  &  $0.000$ \\
                    SmallNORB28x28Inverted         &  $LLR^{>1}$  &  $0.985$  &  $0.987$  &  $0.000$ \\
                    MNIST                          &  $LLR^{>2}$  &  $0.515$  &  $0.507$  &  $0.792$ \\
                    \bottomrule
                \end{tabular}
                }
            \end{table}
        \end{column}
    \end{columns}
}

>>>>

@@

\frame{
    \frametitle{Problem and Contributions}
    \begin{itemize}
        \item Deep generative models often fail at OOD detection task when using their likelihood estimate as the score function \cite{nalisnick_deep_2019} by, perhaps surprisingly, assigning \highlight{higher likelihoods to OOD data}.
        \item Contributions:
        \begin{itemize}
            \item We provide evidence that out-of-distribution detection fails due to learned low-level features that generalize across datasets.
            \item We present a new score for OOD detection with hierarchical VAEs that alleviates this issue.
        \end{itemize}  
    \end{itemize}
}

@@

\begin{frame}
    \frametitle{Out-of-distribution detection with hierarchical VAEs}
    \begin{columns}[t]
        \begin{column}{0.4\textwidth}
            \begin{itemize}
                \item Generative models learn to approximate the \textbf{data distribution} $p(\xb)$.
                \item The likelihood of the model given a sample $\xb$ is a measure of how well the model \textbf{explains the data}.
                \item \textbf{Model likelihood} has long been thought of as useful for OOD detection \cite{bishop_novelty_1994}.
            \end{itemize}
        \end{column}
        \begin{column}{0.6\textwidth}
            \begin{overprint}
                \onslide<2>
                \begin{figure}
                    \centering
                    \includegraphics[width=\textwidth]{../graphics/paper_hierarchical/densities-FashionMNIST-test-MNIST-test-bpp-k-0_sohau.pdf}
                \end{figure}
            \end{overprint}
        \end{column}
    \end{columns}
\end{frame}

@@

\begin{frame}
    \frametitle{Selecting the value of $k$}
    \begin{itemize}
        \item Use validation OOD dataset(s).
        \item Compute $LLR^{>k}$ for different values of $k$ and select the one that maximizes the AUROC.
        \item Compute feature correlations for different values of $k$ and select $k$ at the drop.
    \end{itemize}
    \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{../graphics/paper_hierarchical/feature-correlation-heatmap2.pdf}
    \end{figure}
\end{frame}

@@

\frame{
    \frametitle{Results on diverse datasets}
    \begin{columns}
        \begin{column}{0.45\textwidth}
            \hfill
            \begin{table}[t]
                \centering
                \resizebox{0.9\textwidth}{!}{%
                \begin{tabular}{llrrr}
                    \toprule
                    OOD dataset & Metric & AUROC$\uparrow$ & AUPRC$\uparrow$ & FPR80$\downarrow$ \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on CIFAR10}} \\
                    \midrule
                    SVHN          &  $LLR^{>2}$  &  0.811  &  0.837  &  0.394  \\
                    CIFAR10       &  $LLR^{>1}$  &  0.469  &  0.479  &  0.835  \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on SVHN}} \\
                    \midrule
                    CIFAR10            &  $LLR^{>1}$       &  $0.939$  &  $0.950$  &  $0.052$  \\
                    SVHN               &  $LLR^{>1}$       &  $0.489$  &  $0.484$  &  $0.799$  \\
                    \bottomrule
                \end{tabular}
                }
            \end{table}
            \hfill
        \end{column}
        \begin{column}{0.55\textwidth}
            \hfill
            \begin{table}[t]
                \centering
                \resizebox{0.9\textwidth}{!}{%
                \begin{tabular}{llrrr}
                    \toprule
                    OOD dataset & Metric & AUROC$\uparrow$ & AUPRC$\uparrow$ & FPR80$\downarrow$ \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on FashionMNIST}} \\
                    \midrule
                    MNIST                    & $LLR^{>1}$             &  0.986  &  0.987  &  0.011 \\
                    notMNIST                 &  $LLR^{>1}$            &  0.998  &  0.998  &  0.000 \\
                    KMNIST                   &  $LLR^{>1}$            &  0.974  &  0.977  &  0.017 \\
                    Omniglot28x28            &  $LLR^{>2}$            &  1.000  &  1.000  &  0.000 \\
                    Omniglot28x28Inverted    &  $LLR^{>1}$            &  0.954  &  0.954  &  0.050 \\
                    SmallNORB28x28           &  $LLR^{>2}$            &  0.999  &  0.999  &  0.002 \\
                    SmallNORB28x28Inverted   &  $LLR^{>2}$            &  0.941  &  0.946  &  0.069 \\
   {\color{black!60}FashionMNIST}             &  $LLR^{>1}$            &  0.488  &  0.496  &  0.811 \\
                    \midrule
                    \multicolumn{5}{c}{\textbf{Trained on MNIST}} \\
                    \midrule
                    FashionMNIST                   &  $LLR^{>1}$  &  $0.999$  &  $0.999$  &  $0.000$ \\
                    notMNIST                       &  $LLR^{>1}$  &  $1.000$  &  $0.999$  &  $0.000$ \\
                    KMNIST                         &  $LLR^{>1}$  &  $0.999$  &  $0.999$  &  $0.000$ \\
                    Omniglot28x28                  &  $LLR^{>1}$  &  $1.000$  &  $1.000$  &  $0.000$ \\
                    Omniglot28x28Inverted          &  $LLR^{>1}$  &  $0.944$  &  $0.953$  &  $0.057$ \\
                    SmallNORB28x28                 &  $LLR^{>1}$  &  $1.000$  &  $1.000$  &  $0.000$ \\
                    SmallNORB28x28Inverted         &  $LLR^{>1}$  &  $0.985$  &  $0.987$  &  $0.000$ \\
   {\color{black!60}MNIST}                          &  $LLR^{>2}$  &  $0.515$  &  $0.507$  &  $0.792$ \\
                    \bottomrule
                \end{tabular}
                }
            \end{table}
            \hfill
        \end{column}
    \end{columns}
}

\begin{frame}
    \frametitle{Conclusions}
    \begin{itemize}
        \item Key observations: 
        \begin{itemize}
            \item The likelihood of a generative model is not a good score for OOD detection \cite{nalisnick_deep_2019}.
            \item Strong correlations between some latent variables for different datasets.
            \item Reconstructions of OOD data are good when using full approximate posterior.
        \end{itemize}
        \item[\highlight{$\rightarrow$}] Proposed a new score, $LLR^{>k}$, that uses the conditional prior for the top-most latent variables in the hierarchy.
    \end{itemize}
\end{frame}
